---

title: "Documentation"
output:
  html_document: {toc: true, toc_float: true,
  toc_depth: 2}

---

# PEP project

## Map of stations

```{r, message=FALSE, echo=FALSE, warning=FALSE}
 library(sf)
 library(SWMPr)
 library(tidyverse)
 library(ggrepel)
 library(raster)
 #install.packages(rgdal)
 library(rgdal)


 grb.codes <- SWMPr::site_codes() %>%
   dplyr::filter(nerr_site_id == c("grb"))
 nar.codes <- SWMPr::site_codes() %>%
   dplyr::filter(nerr_site_id == c("nar"))
 wel.codes <- SWMPr::site_codes() %>%
   dplyr::filter(nerr_site_id == c("wel"))
 wqb.codes <- SWMPr::site_codes() %>%
   dplyr::filter(nerr_site_id == c("wqb"))

 codes <- rbind(grb.codes, nar.codes, wel.codes, wqb.codes) %>%
   dplyr::mutate(latitude = as.numeric(latitude),#
                  longitude = as.numeric(longitude)*-1)
 
 site_labels <- codes %>%
   group_by(reserve_name) %>%
   slice(1)

 crs <- "+proj=longlat +lat_1=35 +lat_2=45 +lat_0=40 +lon_0=-77+x_0=0
 +y_0=0 +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
 coast_shp <- rgdal::readOGR(file.path(here::here("data-raw","coast", "USA.shp")), verbose = F)
 crs(coast_shp) <- crs
 coast_sf <- as(coast_shp, "sf")

 xmin = -72
 xmax = -69
 ymin = 41
 ymax = 44
 xlims <- c(xmin, xmax)
 ylims <- c(ymin, ymax)
 crs <- "+proj=longlat +lat_1=35 +lat_2=45 +lat_0=40 +lon_0=-77 +x_0=0 +y_0=0 +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
 m <- ggplot()+
   ggplot2::geom_sf(data = coast_sf) +
   ggplot2::coord_sf(crs = crs, xlim = xlims, ylim = ylims)+
   geom_point(aes(longitude, latitude), data = site_labels, color = "midnightblue", size = 2)+
   geom_label_repel(data = site_labels, aes( longitude, latitude, label = reserve_name),
                    size = 3)
 xbreaks <- seq(-72, -69, 1)
 ybreaks <- pretty(codes$latitude)
 xlabels <- paste0(abs(xbreaks), "°W")
 ylabels <- paste0(abs(ybreaks), "°N")
 m + scale_x_continuous(breaks = xbreaks, labels = xlabels)+
   scale_y_continuous(breaks = ybreaks, labels = ylabels)+
      labs(x = NULL, y = NULL, title = "NERRs - New England Sites")
```

## Narraganset (Daily) {.tabset} 

### Dissolved Oygen
For milligrams per liter:

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
obj <- read.csv(here::here("data/daily_wq.csv"))
nartsdaily <- obj %>% filter(station == "NAR") %>% 
  mutate(date = as.Date(date))

nartsdaily %>% 
filter(Variable == "do_mgl") %>% filter(year >= 2007) %>% 
ggplot(aes(x = date, y = mean_daily))+
geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
geom_line(color = "steelblue")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (mg/L)") +
scale_x_date(date_labels = "%y", date_breaks = "1 year")+
ecodata::theme_ts()

nartsdaily %>% 
filter(Variable == "do_mgl") %>% filter(year >= 2020) %>% 
ggplot(aes(x = date, y = mean_daily))+
geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
geom_line(color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "2020 Data") +
ecodata::theme_ts()
```

For percent saturation:

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
obj <- read.csv(here::here("data/daily_wq.csv"))
nartsdaily <- obj %>% filter(station == "NAR") %>% 
mutate(date = as.Date(date))

nartsdaily %>% filter(Variable == "do_pct") %>% filter(year >= 2007) %>%
ggplot(aes(x = date, y = mean_daily))+
geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue")+
geom_line(color = "steelblue")+
labs(x = "Year", y = "Dissolved Oxygen Concentration (% saturation)")+
scale_x_date(date_labels = "%y", date_breaks = "1 year")+
ecodata::theme_ts()

nartsdaily %>% filter(Variable == "do_pct") %>% filter(year >= 2020) %>%
ggplot(aes(x = date, y = mean_daily))+
geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue")+
geom_line(color = "steelblue")+
scale_x_date(date_labels = "%b", date_breaks = "1 month")+
labs(x = "", y = "Dissolved Oxygen Concentration (% saturation)", title = "2020 Data")+
ecodata::theme_ts()
```

### Temperature

```{r, message=FALSE, echo=FALSE, warning=FALSE}
nartsdaily %>% filter(Variable == "temp") %>% filter(year >= 2007) %>%
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  labs(x = "Year", y = "Temperature (°C)")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  ecodata::theme_ts()

nartsdaily %>% filter(Variable == "temp") %>% filter(year >= 2020) %>%
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Temperature (°C)", title = "2020 Data")+
  ecodata::theme_ts()

```

### Salinity 
```{r, message=FALSE, echo=FALSE, warning=FALSE}
nartsdaily %>% 
  filter(Variable == "sal") %>% filter(year >= 2007) %>% 
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  labs(x = "Year", y = "Salinity Concentration (psu)", title = "") +
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  ecodata::theme_ts()

nartsdaily %>% 
  filter(Variable == "sal") %>% filter(year >= 2020) %>% 
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  labs(x = "Year", y = "Salinity Concentration (psu)", title = "2020 Data") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  ecodata::theme_ts()
```

### Climatology

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(SWMPr)
narts <- all_params_dtrng("nartsnut", c('01/01/2010', '12/31/2020'))
narts2 <- narts %>%  mutate_if(is.character,as.numeric) %>% 
  tidyr::pivot_longer(!datetimestamp, names_to = "Variable", values_to = "Value") %>% 
  tidyr::drop_na(Value) %>% 
  mutate(station = c("NAR")) %>% mutate(year =)

nar<- narts2

nar2 <- nar %>% mutate(month = lubridate::month(datetimestamp, label =TRUE, abbr=TRUE)) %>% 
  group_by(month, Variable, station) %>% summarise(Climatology = mean(Value),
  clim.sd = sd(Value)) %>% ungroup()

only2020data <- nar %>% mutate(month = lubridate::month(datetimestamp, label =TRUE, abbr=TRUE),
                         year = lubridate::year(datetimestamp)) %>% filter(Variable == "chla_n" & year == 2020)

nar2 %>% dplyr::filter(nar2$Variable == "chla_n") %>% 
  ggplot(aes(x = month, y = Climatology, group_by("1")))+
  ecodata::theme_ts()+
  geom_point()+
  geom_point(data = only2020data, aes(x = month, y = Value), color = "red", size = 2)+
  #geom_line(data = only2020data, aes(x = month, y = Value), color = "red")+
  geom_line()+
  geom_errorbar(aes(ymin=Climatology-clim.sd, ymax=Climatology+clim.sd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  labs(x = element_blank(), y = "Chlorophyll (μg/L)")
```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
nar3 <- nar %>% mutate(month = lubridate::month(datetimestamp),
                       year = lubridate::year(datetimestamp)) %>%
  group_by(month, Variable, station) %>% mutate(Climatology = mean(Value)) %>%
  mutate(anom = exp(log(Value)-log(Climatology))) %>% ungroup()

nar3help<- nar3 %>% dplyr::filter(Variable == "chla_n") %>% 
  filter(month == "1") %>% group_by(year, station) %>% 
  summarise(anom.mean = mean(anom)) %>% ungroup()
nar3 %>% dplyr::filter(Variable == "chla_n") %>% filter(month == "1") %>% 
  ggplot(aes(x = year, y = anom))+
  ecodata::theme_ts()+
  geom_point()+
  geom_point(nar3help, mapping = aes(x = year, y = anom.mean), 
             color = "red", shape = 3, size = 3)+
  geom_hline(aes(yintercept = 1), color = "steelblue", size = 2)+
  facet_grid(month~station, scales = "free")+
  labs(x = "Year", y = "Chlorophyll Anomaly (ratio)")+
  ggtitle("")
```

## Great Bay (Daily) {.tabset}

### Dissolved Oxygen 
Milligram per liter:

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
grbdaily <- obj %>% filter(station == "GRB") %>% 
mutate(date = as.Date(date))

grbdaily %>% 
filter(Variable == "do_mgl") %>% filter(year >= 2007) %>% 
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (mg/L")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  ecodata::theme_ts()

grbdaily %>% 
filter(Variable == "do_mgl") %>% filter(year >= 2020) %>% 
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L", title = "2020 Data")+
  ecodata::theme_ts()
```

For percent saturation:

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
obj <- read.csv(here::here("data/daily_wq.csv"))
grbdaily <- obj %>% filter(station == "GRB") %>% 
mutate(date = as.Date(date))

grbdaily %>% filter(Variable == "do_pct") %>% filter(year >= 2007) %>%
ggplot(aes(x = date, y = mean_daily))+
geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue")+
geom_line(color = "steelblue")+
labs(x = "Year", y = "Dissolved Oxygen Concentration (% saturation)")+
scale_x_date(date_labels = "%y", date_breaks = "1 year")+
ecodata::theme_ts()

grbdaily %>% filter(Variable == "do_pct") %>% filter(year >= 2020) %>%
ggplot(aes(x = date, y = mean_daily))+
geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue")+
geom_line(color = "steelblue")+
scale_x_date(date_labels = "%b", date_breaks = "1 month")+
labs(x = "", y = "Dissolved Oxygen Concentration (% saturation)", title = "2020 Data")+
ecodata::theme_ts()
```

### Temperature

```{r, message=FALSE, echo=FALSE, warning=FALSE}
grbdaily %>% filter(Variable == "temp") %>% filter(year >= 2007) %>%
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  labs(x = "Year", y = "Temperature (°C)")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  ecodata::theme_ts()

grbdaily %>% filter(Variable == "temp") %>% filter(year >= 2020) %>%
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Temperature (°C)", title = "2020 Data")+
  ecodata::theme_ts()
```

### Salinity 

```{r, message=FALSE, echo=FALSE, warning=FALSE}
grbdaily %>% 
  filter(Variable == "sal") %>% filter(year >= 2007) %>% 
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  labs(x = "Year", y = "Salinity Concentration (psu)", title = "") +
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  ecodata::theme_ts()

grbdaily %>% 
  filter(Variable == "sal") %>% filter(year >= 2020) %>% 
  ggplot(aes(x = date, y = mean_daily))+
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), color = "lightblue", alpha = 0.75)+
  geom_line(color = "steelblue")+
  labs(x = "Year", y = "Salinity Concentration (psu)", title = "2020 Data") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  ecodata::theme_ts()
```

## Narraganset (Monthly) {.tabset}

### Dissolved Oxygen
For milligrams per liter:

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
monthly <- read_csv(here::here("data/water_quality_pep.csv"))
nartsmonth <- monthly %>% filter(station == "nartswq")
waqmonth <- monthly %>% filter(station == "wqbmpwq")
welmonth <- monthly %>% filter(station == "welinwq")
grbmonth <- monthly %>% filter(station == "grbgbwq") %>% 
mutate(date = as.Date(date))

nartsmonth %>% 
  filter(Variable == "DO_mgl") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (mg/L)")+
  ecodata::theme_ts()

nartsmonth %>% 
  filter(Variable == "DO_mgl") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "2020 data")+
  ecodata::theme_ts()
```

For percent saturation:

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
nartsmonth %>% 
  filter(Variable == "DO_Pct") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (% saturation)") +
  ecodata::theme_ts()

nartsmonth %>% 
  filter(Variable == "DO_Pct") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (% saturation)", title = "2020 data")+
  ecodata::theme_ts()
```

### Temperature 

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
nartsmonth %>% 
  filter(Variable == "Temp") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Temperature (C)")+
  ecodata::theme_ts()

nartsmonth %>% 
  filter(Variable == "Temp") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Temperature (C)", title = "2020 data")+
  ecodata::theme_ts()
```

### Salinity

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
nartsmonth %>% 
  filter(Variable == "Sal") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Salnity (psu)")+
  ecodata::theme_ts()

nartsmonth %>% 
  filter(Variable == "Sal") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Salinity (psu)", title = "2020 data")+
  ecodata::theme_ts()
```

### Chlorophyll 

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
chloro_nutri <- read_csv(here::here("data/nutrient_pep.csv"))
nartsnut <- chloro_nutri %>% filter(station == "nartsnut")
welnut <- chloro_nutri %>% filter(station == "welinnut")
grbnut <- chloro_nutri %>% filter(station == "grbgbnut") %>%
  mutate(date = as.Date(date))

nartsnut%>% 
  filter(Variable == "CHLA_N") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Chlorophyll concentration (µg/L)", title = "")+
  ecodata::theme_ts()
```

### Climatologies

```{r, message=FALSE, echo=FALSE, warning=FALSE}
narclim <- monthly %>% mutate(month = lubridate::month(date, label = TRUE, abbr = TRUE),
                              year = lubridate::year(date)) %>% filter(station == "nartswq") %>% ungroup()

allyearnarclim <- narclim %>% 
group_by(month, Variable, station, year) %>% 
summarise(meanmonth = mean(mean_daily),
          meansd = sd(sd_daily)) %>% filter(year == "2020")

onlymonthnarclim <- narclim %>% 
  group_by(month, Variable, station) %>% 
  summarise(meanmonth = mean(mean_daily),
            meansd = sd(sd_daily))

onlymonthnarclim %>% 
  dplyr::filter(Variable == "DO_mgl") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearnarclim %>% filter(Variable == "DO_mgl"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "Narraganset DO (mg/L) - Climatology")+
  ecodata::theme_ts()

onlymonthnarclim %>% 
  dplyr::filter(Variable == "DO_Pct") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearnarclim %>% filter(Variable == "DO_Pct"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (%)", title = "Narraganset DO (%) - Climatology")+
  ecodata::theme_ts()

onlymonthnarclim %>% 
  dplyr::filter(Variable == "Sal") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearnarclim %>% filter(Variable == "Sal"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Salinity (psu)", title = "Narraganset Salinity - Climatology")+
  ecodata::theme_ts()

onlymonthnarclim %>% 
  dplyr::filter(Variable == "Temp") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearnarclim %>% filter(Variable == "Temp"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Temperature (C)", title = "Narraganset Temperature - Climatology")+
  ecodata::theme_ts()

```

## Great Bay (Monthly) {.tabset}

### Dissolved Oxygen

For milligrams per liter:

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
grbmonth %>% 
  filter(Variable == "DO_mgl") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (mg/L)")+
  ecodata::theme_ts()

grbmonth %>% 
  filter(Variable == "DO_mgl") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "2020 data")+
  ecodata::theme_ts()
```
  
For percent saturation:

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
grbmonth %>% 
  filter(Variable == "DO_Pct") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (% saturation)") +
  ecodata::theme_ts()

grbmonth %>% 
  filter(Variable == "DO_Pct") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (% saturation)", title = "2020 data")+
  ecodata::theme_ts()
```
  
### Temperature

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
grbmonth %>% 
  filter(Variable == "Temp") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Temperature (C)")+
  ecodata::theme_ts()

grbmonth %>% 
  filter(Variable == "Temp") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Temperature (C)", title = "2020 data")+
  ecodata::theme_ts()
```

### Salinity

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
grbmonth %>% 
  filter(Variable == "Sal") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Salnity (psu)")+
  ecodata::theme_ts()

grbmonth %>% 
  filter(Variable == "Sal") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Salinity (psu)", title = "2020 data")+
  ecodata::theme_ts()
```

### Chlorophyll 

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
grbnut%>% 
  filter(Variable == "CHLA_N") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Chlorophyll concentration (µg/L)", title = "")+
  ecodata::theme_ts()
```

### Climatologies

```{r, message=FALSE, echo=FALSE, warning=FALSE}
grbclim <- monthly %>% mutate(month = lubridate::month(date, label = TRUE, abbr = TRUE),
                              year = lubridate::year(date)) %>% filter(station == "grbgbwq") %>% ungroup()

allyeargrbclim <- grbclim %>% 
  group_by(month, Variable, station, year) %>% 
  summarise(meanmonth = mean(mean_daily),
            meansd = sd(sd_daily)) %>% filter(year == "2020")

onlymonthgrbclim <- grbclim %>% 
  group_by(month, Variable, station) %>% 
  summarise(meanmonth = mean(mean_daily),
            meansd = sd(sd_daily))

onlymonthgrbclim %>% 
  dplyr::filter(Variable == "DO_mgl") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyeargrbclim %>% filter(Variable == "DO_mgl"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "Great Bay DO (mg/L) - Climatology")+
  ecodata::theme_ts()

onlymonthgrbclim %>% 
  dplyr::filter(Variable == "DO_Pct") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyeargrbclim %>% filter(Variable == "DO_Pct"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (%)", title = "Great Bay DO (%) - Climatology")+
  ecodata::theme_ts()

onlymonthgrbclim %>% 
  dplyr::filter(Variable == "Sal") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyeargrbclim %>% filter(Variable == "Sal"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Salinity (psu)", title = "Great Bay Salinity - Climatology")+
  ecodata::theme_ts()

onlymonthgrbclim %>% 
  dplyr::filter(Variable == "Temp") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyeargrbclim %>% filter(Variable == "Temp"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Temperature (C)", title = "Great Bay Temperature - Climatology")+
  ecodata::theme_ts()
```

## Waquoit (Monthly) {.tabset}

### Dissolved Oxygen
For milligrams per liter: 

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
waqmonth %>% 
  filter(Variable == "DO_mgl") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (mg/L)")+
  ecodata::theme_ts()

waqmonth %>% 
  filter(Variable == "DO_mgl") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "2020 data")+
  ecodata::theme_ts()
```

For percent saturation:

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
waqmonth %>% 
  filter(Variable == "DO_Pct") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (% saturation)") +
  ecodata::theme_ts()

waqmonth %>% 
  filter(Variable == "DO_Pct") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (% saturation)", title = "2020 data")+
  ecodata::theme_ts()
```

### Temperature 

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
waqmonth %>% 
  filter(Variable == "Temp") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Temperature (C)")+
  ecodata::theme_ts()

waqmonth %>% 
  filter(Variable == "Temp") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Temperature (C)", title = "2020 data")+
  ecodata::theme_ts()
```

### Salinity 

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
waqmonth %>% 
  filter(Variable == "Sal") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Salnity (psu)")+
  ecodata::theme_ts()

waqmonth %>% 
  filter(Variable == "Sal") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Salinity (psu)", title = "2020 data")+
  ecodata::theme_ts()
```

### Climatologies 

```{r, message=FALSE, echo=FALSE, warning=FALSE}
waqclim <- monthly %>% mutate(month = lubridate::month(date, label = TRUE, abbr = TRUE),
                              year = lubridate::year(date)) %>% filter(station == "wqbmpwq") %>% ungroup()

allyearwaqclim <- waqclim %>% 
  group_by(month, Variable, station, year) %>% 
  summarise(meanmonth = mean(mean_daily),
            meansd = sd(sd_daily)) %>% filter(year == "2020")

onlymonthwaqclim <- waqclim %>% 
  group_by(month, Variable, station) %>% 
  summarise(meanmonth = mean(mean_daily),
            meansd = sd(sd_daily))

onlymonthwaqclim %>% 
  dplyr::filter(Variable == "DO_mgl") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwaqclim %>% filter(Variable == "DO_mgl"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "Waquoit DO (mg/L) - Climatology")+
  ecodata::theme_ts()

onlymonthwaqclim %>% 
  dplyr::filter(Variable == "DO_Pct") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwaqclim %>% filter(Variable == "DO_Pct"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (%)", title = "Waquoit Bay DO (%) - Climatology")+
  ecodata::theme_ts()

onlymonthwaqclim %>% 
  dplyr::filter(Variable == "Sal") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwaqclim %>% filter(Variable == "Sal"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Salinity (psu)", title = "Waquoit Bay Salinity - Climatology")+
  ecodata::theme_ts()

onlymonthwaqclim %>% 
  dplyr::filter(Variable == "Temp") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwaqclim %>% filter(Variable == "Temp"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Temperature (C)", title = "Waquoit Bay Temperature - Climatology")+
  ecodata::theme_ts()
```

## Wells (Monthly) {.tabset}

### Dissolved Oxygen

For milligrams per liter:

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
welmonth %>% 
  filter(Variable == "DO_mgl") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (mg/L)")+
  ecodata::theme_ts()

welmonth %>% 
  filter(Variable == "DO_mgl") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "2020 data")+
  ecodata::theme_ts()
```
  
For percent saturation:

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
welmonth %>% 
  filter(Variable == "DO_Pct") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Dissolved Oxygen Concentration (% saturation)") +
  ecodata::theme_ts()

welmonth %>% 
  filter(Variable == "DO_Pct") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Dissolved Oxygen Concentration (% saturation)", title = "2020 data")+
  ecodata::theme_ts()
```
  
### Temperature

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
welmonth %>% 
  filter(Variable == "Temp") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Temperature (C)")+
  ecodata::theme_ts()

welmonth %>% 
  filter(Variable == "Temp") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Temperature (C)", title = "2020 data")+
  ecodata::theme_ts()
```

### Salinity

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
welmonth %>% 
  filter(Variable == "Sal") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Salnity (psu)")+
  ecodata::theme_ts()

welmonth %>% 
  filter(Variable == "Sal") %>% filter(date >= "2020-01-01") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  labs(x = "", y = "Salinity (psu)", title = "2020 data")+
  ecodata::theme_ts()
```

### Chlorophyll

``` {r, message=FALSE, echo=FALSE, warning=FALSE}
welnut%>% 
  filter(Variable == "CHLA_N") %>% 
  ggplot(aes(date, mean_daily))+
  geom_point(color = "lightblue")+
  geom_smooth(method = "loess", span = 0.05, color = "steelblue")+
  scale_x_date(date_labels = "%y", date_breaks = "1 year")+
  labs(x = "Year", y = "Chlorophyll concentration (µg/L)", title = "Wells Chlorophyll Concentration")+
  ecodata::theme_ts()
```

### Climatologies 

```{r, message=FALSE, echo=FALSE, warning=FALSE}
welclim <- monthly %>% mutate(month = lubridate::month(date, label = TRUE, abbr = TRUE),
                              year = lubridate::year(date)) %>% filter(station == "welinwq") %>% ungroup()

allyearwelclim <- welclim %>% 
  group_by(month, Variable, station, year) %>% 
  summarise(meanmonth = mean(mean_daily),
            meansd = sd(sd_daily)) %>% filter(year == "2020")

onlymonthwelclim <- welclim %>% 
  group_by(month, Variable, station) %>% 
  summarise(meanmonth = mean(mean_daily),
            meansd = sd(sd_daily))

onlymonthwelclim %>% 
  dplyr::filter(Variable == "DO_mgl") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwelclim %>% filter(Variable == "DO_mgl"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (mg/L)", title = "Wells Bay DO (mg/L) - Climatology")+
  ecodata::theme_ts()

onlymonthwelclim %>% 
  dplyr::filter(Variable == "DO_Pct") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwelclim %>% filter(Variable == "DO_Pct"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Dissolved Oxygen Concentration (%)", title = "Wells Bay DO (%) - Climatology")+
  ecodata::theme_ts()

onlymonthwelclim %>% 
  dplyr::filter(Variable == "Sal") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwelclim %>% filter(Variable == "Sal"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Salinity (psu)", title = "Wells Bay Salinity - Climatology")+
  ecodata::theme_ts()

onlymonthwelclim %>% 
  dplyr::filter(Variable == "Temp") %>% 
  ggplot(aes(month, meanmonth))+
  geom_point(color = "lightblue")+
  geom_errorbar(aes(ymin=meanmonth-meansd, ymax=meanmonth+meansd), width=.2,
                position=position_dodge(0.05), color = "steelblue")+
  geom_point(data = allyearwelclim %>% filter(Variable == "Temp"), aes(month, meanmonth), color = "red", size = 2)+
  labs(x = "", y = "Temperature (C)", title = "Wells Bay Temperature - Climatology")+
  ecodata::theme_ts()
```
