---
title: 'GAM meta analysis data exploration'
author: "Kimberly Bastille"
date: "6/22/2022"
output: html_document
---

### is there enough data?

Below is a summary of the data comparisons between the Beck et. al 2022 paper and the NERRS data. 

| Parameter |  Beck et. al 2022 | NERRS data from stations used in Rhegan's Analysis |
|---------------|------------------|-------------------------------------------------|
| Time Series Length | 30 years (1990 - 2019) | 15 years (2007 - 2019) * earlier samples removed due to poor qaqc scores |
| Number of total samples | 414 - 608 samples per station | 34 - 470 *greatest from Wells Bay |
| Seasons available | ALL | Summer *due to ice cover summer was the season most likely to show trends with enough data |

Summary tables describing data from New England NERRS stations. (welinnut = Wells Bay ME, grbgbnut = Great Bay NH, wqbmpnut = Waquoit Bay MA, nartsnut = Narragansett RI) 

```{r, echo=FALSE,  message=FALSE, warning=FALSE}
library(tidyverse)
library(SWMPr)
raw_dat <- bind_rows(
  import_local(here::here("data-raw/nut"), station_code = "nartsnut") %>% 
    qaqc() %>% 
    mutate(station = "nartsnut"),
  import_local(here::here("data-raw/nut"), station_code = "wqbmpnut") %>% 
    qaqc() %>% 
    mutate(station = "wqbmpnut"),
  import_local(here::here("data-raw/nut"), station_code = "welinnut") %>%
    qaqc() %>%
    mutate(station = "welinnut"),
  import_local(here::here("data-raw/nut"), station_code = "grbgbnut") %>%
    qaqc() %>%
    mutate(station = "grbgbnut"))
```

#### Station Data Summary
```{r, echo=FALSE, message=FALSE, warning=FALSE}
total_nut <- raw_dat %>% 
  mutate(date = format(as.Date(datetimestamp), "%Y-%m-%d")) %>% 
  select(date,
         station,
         #po4f = po4f, 
         #nh4f = nh4f, 
         #no2f = no2f,
         #no23f = no23f, 
         chla_n = chla_n) %>% #, %>% 
         #no3f = no3f) %>%
  #pivot_longer(cols = c(chla_n), 
  #             names_to = "Variable", values_to = "value") %>% 
  dplyr::mutate(month = lubridate::month(date), 
                year = lubridate::year(date))%>%
  dplyr::filter(month > 5, 
                month < 9) %>% 
  group_by(station) %>%
  dplyr::filter(!chla_n == "NA") %>% 
  summarise(Number_of_Stations = length(chla_n), 
            Years_Missing = c())
years_missing <- data.frame(station = c("grbgbnut", "nartsnut", "welinnut", "wqbmpnut"),
                            Years_Missing = c("2009,2010,2014,2015,2016,2017", 
                                             "2008.2009,2010", 
                                             "2016", 
                                             "2014"))
dat_tot <- total_nut %>%  left_join(years_missing)
knitr::kable(dat_tot)

```

#### Seasonal Data Summary

```{r, echo=FALSE, message=FALSE, warning=FALSE}


seasonal_nut <- raw_dat %>% 
  mutate(date = format(as.Date(datetimestamp), "%Y-%m-%d")) %>% 
  select(date,
         station,
         #po4f = po4f, 
         #nh4f = nh4f, 
         #no2f = no2f,
         #no23f = no23f, 
         chla_n = chla_n) %>% #, %>% 
         #no3f = no3f) %>%
  #pivot_longer(cols = c(chla_n), 
  #             names_to = "Variable", values_to = "value") %>% 
  dplyr::mutate(month = lubridate::month(date), 
                year = lubridate::year(date))%>%
  dplyr::filter(month > 5, 
                month < 9) %>% 
  group_by(station, year) %>%
  dplyr::filter(!chla_n == "NA") %>% 
  summarize(N_stations = length(chla_n),
            mean_seasonal = mean(chla_n), 
            se_seasonal = sd(chla_n)/sqrt(N_stations)) 

seasonal_nut %>% 
  ggplot(aes(x = year, y = mean_seasonal, group = station, color = station))+
  geom_point()+
  geom_path()+
  geom_errorbar(aes(ymin=mean_seasonal-se_seasonal, ymax=mean_seasonal+se_seasonal), width=.2)+
  ggtitle("Chlorophyll A Summer Means for New England NERRs")

seasonal_nut %>% 
  dplyr::filter(!station == "nartsnut",
                !station == "grbgbnut") %>% 
  ggplot(aes(x = year, y = mean_seasonal))+
  geom_point()+
  geom_path()+
  geom_errorbar(aes(ymin=mean_seasonal-se_seasonal, ymax=mean_seasonal+se_seasonal), width=.2)+
  facet_grid(rows = vars(station), scales = "free")+
  ggtitle("Chlorophyll A Summer Means for Wells (data rich) and Waquoit (data poor)")

DT::datatable(seasonal_nut)

```

### GAM meta analysis
```{r, echo=FALSE, message=FALSE, warning=FALSE}


```


